---
title: "Lab9 Base R - No Packages"
author: "Sage Davis"
date: "11/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objects and Classes

## Vectors

### Creating a Vector

```{r}
# numeric vector
my_vector <- vector(mode = "numeric", length = 100)

# logical
my_second_vector <- vector(mode = "logical", length = 20)

```

### Vector Properties

#### Length

```{r}
length(my_vector)

length(my_second_vector)

```

#### Class

```{r}

class(my_vector)

class(my_second_vector)

```

#### Indexing (accessing) vectors

```{r}

# using c to create a vector
m_vector <- c(1:50)

# what is the 1st element
m_vector[1]

# what is the last element
m_vector[-1]

# what is the 7th element
m_vector[7]

# select a range from a vector
m_vector[3:7]
```

## Matricies

### Creating Metricies

```{r}
# Create a 10 x 10 vector of NAs
my_matrix <- matrix(data = 1:100, nrow = 20, ncol = 5)
```

### Matic Properties

#### Dimensions

```{r}
# What are the dimensions
dim(my_matrix)
```

#### Length

```{r}
# How many elements are there?
length(my_matrix)
```

#### Class

```{r}
class(my_matrix)
```


#### Indexing

```{r}
my_matrix[2,]  # 1st numer is always row

my_matrix[,3]  # Index a column

my_matrix[8,3]  # index a cell

my_matrix[8,3] <- 100000  # replace a cell

my_matrix[,5] <- 10   # replace a column
```


# Control Flow

## 'for' loops 

variable (counter), sequence, expression - here - i, 1:10, print(i)

```{r}
for (i in 1:10){
  print(i)  # replce i with "Hello World", will print 10x
}
```

# Gemerating random numbers

```{r}
rnorm(n = 5, mean = 10, sd = 2)
```

# Let's combine all that we have learned

We will use 'numeric' to create an empty vector
we will use 'for' to iterate along items of that empty vector
we will use '[]' to access each slot and fill-in that empty vector
we will use 'rnorm' to sample values from the random normal distr
we will plot the trend
# $$ is an equation $$

$$
N_{t}=N_{t-1} + r\left{1 - frac{N_{t-1}}{K}\right}
$$

$$
r \sim norm(\mu = 1.5, \sigma = 0.8)
$$
```{r}
# set up the model parameters
K <- 100
nsteps <- 100
mu <- 0.4
sigma <- 1

# set up model objets
N <- vector(mode = "numeric", length = nsteps)
N[1] <- 25

# Model it
for(t in 2:nsteps){
  r <- rnorm(n = 1, mean = mu, sd = sigma)
  N[t] <- N[t-1] + r * (1 - N[t-1] / K)
}

# Visualize
plot(N)
```
# Something More

We will use 'matrix' to create an empty matrix (rows are time, col are runs)
We will repeat the process


```{r}
# set up the model parameters
K <- 100
nsteps <- 100
mu <- 0.4
sigma <- 1

# set up model objets
N <- matrix(data = NA, nrow = nsteps, ncol = 100)
N[1,] <- 25

# Model it
for(t in 2:nsteps){
  r <- rnorm(n = 100, mean = mu, sd = sigma)
  N[t,] <- N[t-1,] + r * (1 - N[t-1,] / K)
}

# Visualize
matplot(N, type = "l")
```

# Bring in your friends

Convert it to things we know about: 'ggplot'

```{r}
library(tidyverse)

# create data.frame
my_data <- as_tibble(N) %>% 
  mutate(time = 1:nsteps) %>% 
  gather(run, N, -time)

# visualize it
ggplot(data = my_data, mapping = aes(x = time, y = N)) + 
  geom_line(aes(group = run), size = 0.1, alpha = 0.2) +
  stat_summary(geom = "line", fun = mean, color = "red") +
  theme_bw() +
  labs(x = "Time (years)", y = "Population size")

```


